CS6213 Project Report

### 2021-05-07 report/paper.tex
Pain points

\section{TLA Specification}

@@

    \item \textbf{Views} provide a way to change when TLC considers two states to be equal.
    %
    Two states are equal if their view evaluates to to the same value.
    %
    We use a view to instruct TLC to ignore history variables when comparing states.
    %
    This leads to a hundredfold decrease in number of states explored.

>>>>

\section{TLA}

\subsection{Pain Points}

Our experience with TLA was somewhat frustrating.

\subsubsection{Extending Specifications}

In our experience, it is fairly easy to specify complex behaviours when you are writing the specification from scratch, but it is a hassle to \emph{extend} an existing spec, \eg, to add history variables to Raft. 
%
TLA specifications are very non-compositional, and even small extensions can require non-local changes.

In our case, we had a variable called \texttt{history}, which already recorded some history information, and we wanted additionally to make this variable record when messages were sent.
%
Intuitively, you would expect that changing the definition of \texttt{Send} to assign \texttt{history'} would be enough, but in fact, in TLA, we had to also change every operator that directly or indirectly depended on \texttt{Send} to remove the \texttt{UNCHANGED <<history>>} clause.

We had a similar experience when we tried to define a \texttt{Broadcast} operator in terms of \texttt{Send}.
%
We gave up and defined it from scratch.

\subsubsection{Runtime Errors}

On several ocassions, we encountered (trivial) runtime errors after considerable model-checking time (> 15 minutes).
%
We know this is an issue in practice, and not just for us, because the specification of one-at-a-time membership changes for Raft that we adapted had two such issues which, with the default TLC options, would have taken on the order of hours to manifest at runtime.
%
This class of problem would be solved by the addition of a type system to TLA.

@@

    \item \textbf{Views} provide a way to change when TLC considers two states to be equal.
    %
    Two states are equal if their view evaluates to to the same value.
    %
    We use a view to instruct TLC to ignore history variables when comparing states.
    %
    This leads to a hundredfold decrease in number of states explored.
\end{itemize}

\subsection{Concolic Trace Generation}
